### 2024-04-19 docs/SOSP 2024/schedule.tex
pass

\section{Scheduling Language}
\label{sec:schedule}
% As established in Section \ref{sec:motivation}, there are several 
% configurations and optimizations strategies for decision 
% tree inference. The best one to use varies significantly across
% models, batch sizes and hardware platforms. Therefore, designing
% any one hard-coded strategy is not feasible as this would 
% make portable performance impossible. 
This section discusses the scheduling language that enables exploration of different strategies for decision tree inference. The language provides an abstract way to specify loop structure and 
other optimizations as an input to the compiler. The specified 
\textbf{\emph{schedule}} controls the lowering of model inference to a set of 
loop nests. The configurability provided by the schedule allows us 
to build auto-schedulers and auto-tuners (Section \ref{sec:exploring}).
% Using a scheduling language will also significantly simplify 
% adding support for new hardware as this will likely require
% different locality optimizations and loop transformations.

@@

  \bottomrule
  \end{tabularx}
  }
  \vskip 5pt
  \caption{\label{Tab:Optimizations} List of optimization directives in \Treebeard{}'s scheduling language. 
  We use \emph{index variable} and \emph{loop} interchangeably in descriptions for clarity of exposition.}
\end{table}

@@

  \bottomrule
  \end{tabularx}
  }
  \vskip 5pt
  \caption{\label{Tab:ReductionOpts} List of reduction optimization directives in \Treebeard{}'s scheduling language. 
  We use \emph{index variable} and \emph{loop} interchangeably in descriptions for clarity of exposition.}
\end{table}

@@

XGBoost\cite{XGBoost} implements inference on the CPU by going 
over a fixed number of rows (64 in the previous version)
for every tree and then moving to the next tree. It moves to the 
next set of rows when all trees have been walked for the current 
set of rows. Different sets of rows are processed in parallel.
This is expressed in \Treebeard{}'s scheduling language as:
\begin{lstlisting}[style=c++]
  tile(batch, b0, b1, CHUNK_SIZE)
  reorder(b0, tree, b1)
  parallel(b0)
\end{lstlisting}

@@

In the \emph{direct method}~\cite{Tahoe}, a single GPU thread walks all trees
for a given input row. The schedule for this strategy is as follows.
\begin{lstlisting}[style=c++]
  tile(batch, b0, b1, ROWS_PER_TB)
  reorder(b0, b1, tree)
  gpuDimension(b0, grid.x)
  gpuDimension(b1, block.x)
\end{lstlisting}
Here, \op{ROWS\_PER\_TB} is the number of rows that are processed by a single thread block.

In the \emph{shared data} strategy~\cite{Tahoe}, a thread block walks all the trees 
for a given row in parallel. 
% If threads walk multiple trees, each thread accumulates partial results. 
Then, a thread block wide reduction is performed to compute 
the prediction. The schedule for this strategy is as follows.
\begin{lstlisting}[style=c++]
reorder(batch, tree)
gpuDimension(batch, grid.x)
gpuDimension(tree, block.x)
cache(batch)
\end{lstlisting}
%   \item \textbf{Shared Forest}: In this strategy, the whole model is loaded into 
%   shared memory and subsequently, a single thread walks all trees for a particular
%   row. The schedule for this strategy is as follows.
% \begin{lstlisting}[style=c++]
%   tile(batch, b0, b1, ROWS_PER_TB)
%   tile(tree, t0, t1, N_TREES)
%   reorder(b0, b1, t0, t1)
%   cache(t0)
%   gpuDimension(b0, grid.x)
%   gpuDimension(b1, block.x)
% \end{lstlisting}
%   Here, we create a placeholder single iteration loop \op{t0} so that we can 
%   specify that all trees are to be cached.
%   \item \textbf{Shared Partial Forest}: In case the model is too large to fit into
%   shared memory, the model is split into chunks and each chunk is loaded into shared
%   memory. Again, as in the previous strategy, one thread walks all trees assigned to
%   a thread block for a row. The schedule for this strategy is as follows.
% \begin{lstlisting}[style=c++]
%   tile(batch, b0, b1, ROWS_PER_TB)

>>>>

\section{Scheduling Language}
\label{sec:schedule}
% As established in Section \ref{sec:motivation}, there are several 
% configurations and optimizations strategies for decision 
% tree inference. The best one to use varies significantly across
% models, batch sizes and hardware platforms. Therefore, designing
% any one hard-coded strategy is not feasible as this would 
% make portable performance impossible. 
This section presents all constructs in the scheduling language. 
The language provides an abstract way to specify loop structure and 
other optimizations as an input to the compiler. The specified 
\textbf{\emph{schedule}} controls the lowering of model inference to a set of 
loop nests. The configurability provided by the schedule allows us 
to build auto-schedulers and auto-tuners (Section \ref{sec:exploring}).
% Using a scheduling language will also significantly simplify 
% adding support for new hardware as this will likely require
% different locality optimizations and loop transformations.

@@

  \bottomrule
  \end{tabularx}
  }
  \vskip 5pt
  \caption{\label{Tab:Optimizations} List of optimization directives in \Treebeard{}'s scheduling language. 
 % We use \emph{index variable} and \emph{loop} interchangeably in descriptions for clarity of exposition.
 }
\end{table}

@@

  \bottomrule
  \end{tabularx}
  }
  \vskip 5pt
  \caption{\label{Tab:ReductionOpts} List of reduction optimization directives in \Treebeard{}'s scheduling language. 
  %We use \emph{index variable} and \emph{loop} interchangeably in descriptions for clarity of exposition.
  }
\end{table}

@@

XGBoost\cite{XGBoost} implements inference on the CPU by going 
over a fixed number of rows (64 in the current version)
for every tree and then moving to the next tree. It moves to the 
next set of rows when all trees have been walked for the current 
set of rows. Different sets of rows are processed in parallel.
This is expressed in \Treebeard{}'s as:
\begin{lstlisting}[style=c++]
  tile(batch, b0, b1, CHUNK_SIZE)
  reorder(b0, tree, b1)
  parallel(b0)
\end{lstlisting}

@@

In the \emph{direct method}~\cite{Tahoe}, a single GPU thread walks all trees
for a given input row. The schedule for this strategy is:
\begin{lstlisting}[style=c++]
  tile(batch, b0, b1, ROWS_PER_TB)
  reorder(b0, b1, tree)
  gpuDimension(b0, grid.x)
  gpuDimension(b1, block.x)
\end{lstlisting}
Here, \op{ROWS\_PER\_TB} is the number of rows that are processed by a single thread block.

In the \emph{shared data} strategy~\cite{Tahoe}, a thread block walks all the trees 
for a given row in parallel. 
% If threads walk multiple trees, each thread accumulates partial results. 
Then, a thread block wide reduction is performed to compute 
the prediction. The schedule for this strategy is:
\begin{lstlisting}[style=c++]
reorder(batch, tree)
gpuDimension(batch, grid.x)
gpuDimension(tree, block.x)
reduce(batch)
\end{lstlisting}
%   \item \textbf{Shared Forest}: In this strategy, the whole model is loaded into 
%   shared memory and subsequently, a single thread walks all trees for a particular
%   row. The schedule for this strategy is as follows.
% \begin{lstlisting}[style=c++]
%   tile(batch, b0, b1, ROWS_PER_TB)
%   tile(tree, t0, t1, N_TREES)
%   reorder(b0, b1, t0, t1)
%   cache(t0)
%   gpuDimension(b0, grid.x)
%   gpuDimension(b1, block.x)
% \end{lstlisting}
%   Here, we create a placeholder single iteration loop \op{t0} so that we can 
%   specify that all trees are to be cached.
%   \item \textbf{Shared Partial Forest}: In case the model is too large to fit into
%   shared memory, the model is split into chunks and each chunk is loaded into shared
%   memory. Again, as in the previous strategy, one thread walks all trees assigned to
%   a thread block for a row. The schedule for this strategy is as follows.
% \begin{lstlisting}[style=c++]
%   tile(batch, b0, b1, ROWS_PER_TB)
